Create a histogram of diamond prices.
Facet the histogram by diamond color
and use cut to color the histogram bars.

The plot should look something like this.
http://i.imgur.com/b5xyrOu.jpg

Note: In the link, a color palette of type
'qual' was used to color the histogram using
scale_fill_brewer(type = 'qual')

This assignment is not graded and
will be marked as correct when you submit.


```{r}

library(ggplot2)
data(diamonds)

ggplot(aes(x=price),
       data = diamonds)+
  facet_wrap(~color)+
  geom_histogram(aes(fill=cut))+
  scale_fill_brewer(type='qual')
```

Create a scatterplot of diamond price vs.
table and color the points by the cut of
the diamond.

The plot should look something like this.
http://i.imgur.com/rQF9jQr.jpg

Note: In the link, a color palette of type
'qual' was used to color the scatterplot using
scale_color_brewer(type = 'qual')
```{r}
str(diamonds)
ggplot(aes(x=table,y=price),
       data = diamonds)+
  geom_point(aes(color=cut))+
  scale_color_brewer(type='qual')
```

The discrete will shows that ideal 53-57, premium 58-62




Create a scatterplot of diamond price vs.
volume (x * y * z) and color the points by
the clarity of diamonds. Use scale on the y-axis
to take the log10 of price. You should also
omit the top 1% of diamond volumes from the plot.

Note: Volume is a very rough approximation of
a diamond's actual volume.

The plot should look something like this.
http://i.imgur.com/excUpea.jpg
```{r}

diamonds <- transform(diamonds,volume=x*y*z)
ggplot(aes(x=volume,y=price),
       data=diamonds)+
  geom_point(aes(color=clarity))+
  scale_y_log10()+
  xlim(0,quantile(diamonds$volume,0.99))+
  scale_color_brewer(type='div')
```


Many interesting variables are derived from two or more others.
For example, we might wonder how much of a person's network on
a service like Facebook the user actively initiated. Two users
with the same degree (or number of friends) might be very
different if one initiated most of those connections on the
service, while the other initiated very few. So it could be
useful to consider this proportion of existing friendships that
the user initiated. This might be a good predictor of how active
a user is compared with their peers, or other traits, such as
personality (i.e., is this person an extrovert?).

Your task is to create a new variable called 'prop_initiated'
in the Pseudo-Facebook data set. The variable should contain
the proportion of friendships that the user initiated.
```{r}
pf <- read.csv('../lesson3/pseudo_facebook.tsv',sep='\t')
names(pf)
pf <- transform(pf, prop_initiated=friendships_initiated/friend_count)
```



Create a line graph of the proportion of
friendships initiated ('prop_initiated') vs.
tenure and color the line segment by
year_joined.bucket.

Recall, we created year_joined.bucket in Lesson 5
by first creating year_joined from the variable tenure.
Then, we used the cut function on year_joined to create
four bins or cohorts of users.

(2004, 2009]
(2009, 2011]
(2011, 2012]
(2012, 2014]
```{r}
pf <- transform(pf,year_joined=floor(2014-tenure/365))
pf <- transform(pf, year_joined.bucket = cut(year_joined,c(2004,2009,2011,2012,2014)))
table(pf$year_joined.bucket)

ggplot(aes(x=tenure,y=prop_initiated),
       data=pf)+
  geom_line(aes(color=year_joined.bucket),
            stat="summary",
            fun.y = median)
```


Smooth the last plot you created of
of prop_initiated vs tenure colored by
year_joined.bucket. You can use larger
bins for tenure or add a smoother to the plot.

There won't be a solution image for this exercise.
You will answer some questions about your plot in
the next two exercises.
```{r}
ggplot(aes(x=tenure,y=prop_initiated),
       data=pf)+
  geom_smooth(aes(color=year_joined.bucket))
summary(pf$year_joined.bucket)
```


```{r}
summary(subset(pf,year_joined.bucket == '(2012,2014]')$prop_initiated)
```

* Because people with late join, tends to have friends that already on facebook. So they're catching up by invite their friends
* When they first join, they were asked to invite friends that they do know, instead earlier when facebook didn't implement such system.
* Facebook has become very famous, and sort of de facto standard for social network. Nonetheless, people (must) use facebook for engaging their network
* Like google, many apps nowadays in smartphone using facebook account to quickly login into their account.

****

Create a scatter plot of the price/carat ratio
of diamonds. The variable x should be
assigned to cut. The points should be colored
by diamond color, and the plot should be
faceted by clarity.

The plot should look something like this.
http://i.imgur.com/YzbWkHT.jpg.

Note: In the link, a color palette of type
'div' was used to color the histogram using
scale_color_brewer(type = 'div')

```{r}

ggplot(aes(x=cut, y = price/carat),
       data=diamonds)+
  geom_jitter(aes(color=color))+
  facet_wrap(~clarity)+
  scale_color_brewer(type='div')
```

Extract only ASEAN data from the dataset
create bucket year ~1900,1900-1990,1990-2000,2000-2005
Create time series for the data.
Mean Age at first mariage of women
```{r}
library(dplyr)
wm <- read.csv('indicator age of marriage.csv')

ASEAN <- c('Indonesia','VietNam', 'Myanmar','Thailand','Singapore',
           'Cambodia','Philippines','Brunei','Malaysia')
bucket_year <- c(1616,1900,1990,2000,2005)
wm_asean <- filter(wm,X %in% ASEAN )
wm_asean <- transform(wm_asean, X=factor(X,ASEAN))
colnames(wm_asean) <- as.numeric(substr(colnames(wm_asean),start=2,stop=5))
colnames(wm_asean)[1] <- 'country'
library(reshape2)
wm_asean.long <- melt(wm_asean,
                      id='country',
                      value.name='Average',
                      variable.name = 'Year',
                      measure.vars=colnames(wm_asean)[-1])
wm_asean.long$Year <- as.numeric(as.character(wm_asean.long$Year))
wm_asean.long$Average[is.na(wm_asean.long$Average)] <- 0
wm_asean.long$year_bucket <- cut(wm_asean.long$Year,
                   breaks= bucket_year,
                   labels=c('less than 1900',
                        '1900-1990',
                        '1990-2000',
                        '2000-2005'))

ggplot(aes(x=Year,y=Average),
       data=wm_asean.long)+
  geom_smooth(aes(color=year_bucket))+
  facet_wrap(~country)+
  ggsave('fmasean.jpg')

```






