
Observation = 53940
variable = 10


```{r}
data(diamonds)
names(diamonds)
names(diamonds['color'])
summary(diamonds)
#?ordered
#?levels
ggplot(aes(x=price), data = diamonds) + geom_histogram()
dim(subset(diamonds, diamonds$price >= 15000))
ggplot(aes(x=price), data = diamonds) + geom_histogram(binwidth = 250)+
  scale_x_continuous(breaks = seq(0,5000,250), limit = c(0,5000))

```

There are no diamonds that cost $1500.

For diamonds that cost less than $2,000, the most common price of a diamond is around $700 with the mode being $605 (binwidth = 1). The second insight comes from Paula F. Thanks Paula!

By Cut
```{r}
ggplot(aes(x=price), data = diamonds) + geom_histogram(binwidth = 250)+
  scale_x_continuous(breaks = seq(0,5000,250), limit = c(0,5000)) + facet_wrap(~cut, ncol=2)
```

Question based on Cut
```{r}
by(diamonds$price,diamonds$cut,summary)

by(diamonds$price,diamonds$cut,max)
```

# In the two last exercises, we looked at
# the distribution for diamonds by cut.

# Run the code below in R Studio to generate
# the histogram as a reminder.

# ===============================================================

```{r}
#qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = "free_y") +
ggplot(aes(x=price*carat), data = na.omit(diamonds)) + geom_histogram(binwidth = 250)+
  facet_wrap(~cut, scales  = "free_y") + scale_x_log10()
```
                                                                               
# ===============================================================


# In the last exercise, we looked at the summary statistics
# for diamond price by cut. If we look at the output table, the
# the median and quartiles are reasonably close to each other.

# diamonds$cut: Fair
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     337    2050    3282    4359    5206   18570 
# ------------------------------------------------------------------------ 
# diamonds$cut: Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     327    1145    3050    3929    5028   18790 
# ------------------------------------------------------------------------ 
# diamonds$cut: Very Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     336     912    2648    3982    5373   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Premium
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326    1046    3185    4584    6296   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Ideal
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326     878    1810    3458    4678   18810 

# This means the distributions should be somewhat similar,
# but the histograms we created don't show that.

# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

# If you want a hint, check out the Instructor Notes.


#price per carat natural log distribution and cut facet
```{r}
#qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = "free_y") +
ggplot(aes(x=price/carat), data = na.omit(diamonds)) + geom_histogram()+
  facet_wrap(~cut, scales  = "free_y") + scale_x_log10()
```

#boxplot categorical
```{r}
ggplot(aes(x = color, y = price/carat), data = na.omit(diamonds)) + geom_boxplot()
```

#IQR
```{r}
IQR(subset(diamonds,price<1000)$price)
names(diamonds)
summary(subset(diamonds,color=='J')$price)
IQR(subset(diamonds,color=='D')$color)
by(diamonds$price, diamonds$color, IQR)
```
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

```{r}
ggplot(aes(x = color, y = price/carat), data = diamonds) + geom_boxplot(aes(fill=color)) +
  ggsave('pricePerCaratInColor.png')
table(diamonds$carat)[table(diamonds$carat)>2000]
dim(diamonds)
subset(diamonds$carat, diamonds$carat>0.3)
```
weight diamonds using freq_poly
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.
```{r}
alco = read.csv('alcohol_indicator.csv', header=TRUE, row.names=1)
?geom_freqpoly
ggplot(data = alco) + 
  geom_freqpoly(aes(color = 'Year',x = alco$X2005, y = (..count..)/sum(..count..)))+ 
  geom_freqpoly(aes(color = 'blue',x = alco$X2008, y = (..count..)/sum(..count..)))+
  xlab('count litre per people in a year')+
  ylab('average drink over countries')+
  scale_colour_manual(values = c('green','blue'),
                      labels = c('2005','2008'))+
  ggsave('alcoholplot.png')
?scale_colour_manual
```

# Your task is to investigate the distribution of your friends'
# birth months and days.

# Here some questions you could answer, and we hope you think of others.

# **********************************************************************







# **********************************************************************
```{r}
library(lubridate)
library(gridExtra)
fbfriends = read.csv('../Lesson3/birthdaysExample.csv')
fbfriends$gf <- mdy(fbfriends$dates)
fbfriends$year <- year(fbfriends$gf)
fbfriends$month <- month(fbfriends$gf)
fbfriends$day <- day(fbfriends$gf)
```
# How many people share your birthday? Do you know them?
# (Reserve time with them or save money to buy them a gift!)
```{r}
nrow(subset(fbfriends, fbfriends$month == 12 & fbfriends$day ==21))
```
# Which month contains the most number of birthdays?
```{r}
tbmonth = table(fbfriends$month)

which(tbmonth == max(tbmonth))
```

# How many birthdays are in each month?
```{r}
table(fbfriends$month)
```
# Which day of the year has the most number of birthdays?
```{r}
tbday <- table(fbfriends$day)
which(tbday== max(tbday))
```
# Do you have at least 365 friends that have birthdays on everyday
# of the year?
No

```{r}
dftbmonth = as.data.frame(tbmonth)
plot(dftbmonth$Var1, dftbmonth$Freq)
ggplot(aes(Var1,Freq), data = dftbmonth) + geom_point()
```

