
setting up library

```{r}
library(ggplot2)
library(gridExtra)
library(dplyr)
```

Scatter Plot between price and x in diamonds data set
```{r}
data(diamonds)
names(diamonds)
ggplot(aes(x=x,y=price),data=subset(diamonds, x > 3 & x < 9))+
  geom_point(alpha = 1/20)
```

Observation
the majority of x is between 3-9
the maximum of price is less than 2000
It seems like monotonic increasing function

Correlation
```{r}
cor.test(diamonds$price,diamonds$x)
cor.test(diamonds$price,diamonds$y)
cor.test(diamonds$price,diamonds$z)
```
scatter plot price vs depth
```{r}
ggplot(aes(x=depth, y = price), data =diamonds)+
  geom_point()
```

Adjustment
```{r}
ggplot(aes(x=depth, y = price), data =diamonds)+
  geom_point(alpha = 1/100)+
  scale_x_continuous(breaks = seq(50,70,2))
```
Correlation
```{r}
cor.test(diamonds$depth, diamonds$price)
```
Scatter plot, omit top 1%
```{r}
p1 <- ggplot(aes(x=carat,y=price),data=diamonds)+
  geom_point(stat= "summary", fun.y = quantile, probs = 0.99)
p2 <- ggplot(aes(x=carat,y=price),data=diamonds)+
  geom_point(stat= "summary", fun.y = quantile, probs = 0.01)
grid.arrange(p1,p2,ncol=1)
```
price vs volume
```{r}
diamonds$volume <- (diamonds$x*diamonds$y)*diamonds$z
ggplot(aes(x=volume,y=price), data=diamonds)+
  geom_point()
```
#Observations
seems like linear increase and we can find highly correlation betwen volume and price
Did you notice some outliers? Some volumes are 0! There's an expensive diamond with a volume near 4000, and a cheaper diamond with a volume near 900.

You can find out how many diamonds have 0 volume by using count(diamonds$volume == 0). The count() function comes with the plyr package.

Note: You need to run this command in R to unload the plyr package. detach("package:plyr", unload=TRUE) The plyr package will conflict with the dplyr package in later exercises.

Depending on your investigation, it may or may not be important for you to understand how outliers, like these, came to be in your data.


Correlation price vs volume
```{r}
subdiammondsbyvol <-subset(diamonds,volume>0 & volume <= 800) 
cor.test(subdiammondsbyvol$price, subdiammondsbyvol$volume)
```
# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)

# We encourage you to think about this next question and
# to post your thoughts in the discussion section.

# Do you think this would be a useful model to estimate
# the price of diamonds? Why or why not?

```{r}
ggplot(aes(x=volume, y=price), data=subdiammondsbyvol)+
  geom_point(alpha=1/20)+
  geom_smooth()+
  ggsave("volpricediamond.jpg")
  #stat_smooth(method="loess",size=1)
  #geom_line(stat="summary", fun.y = mean)
```
Use the function dplyr package
 to create a new data frame containing
info on diamonds by clarity.

 Name the data frame diamondsByClarity

 The data frame should contain the following
 variables in this order.

       (1) mean_price
       (2) median_price
       (3) min_price
      (4) max_price
       (5) n

 where n is the number of diamonds in each
level of clarity.
```{r}
library(dplyr)
diamondsByClarity<-diamonds%.%
  group_by(clarity)%.%
  summarise(mean_price=mean(price),
            median_price=median(price),
            min_price=min(price),
            max_price=max(price),
            n=n())
  
```
#Bar Charts of Mean Price
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x=clarity,y=mean_price),data=diamonds_mp_by_clarity)+
  geom_bar(stat="identity")
p2 <- ggplot(aes(x=color,y=mean_price),data=diamonds_mp_by_color)+
  geom_bar(stat="identity")
grid.arrange(p1,p2,ncol=1)
```
Observation
we can't measure the price by clarity, it's categorical, so we plot in bar chart
As opposite, the the color, can be measured by the color, we plot in histogram

We think something odd is going here. These trends seem to go against our intuition.

Mean price tends to decrease as clarity improves. The same can be said for color.

We encourage you to look into the mean price across cut.
diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```{r}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

ggplot(aes(x=cut,y=mean_price),data=diamonds_mp_by_cut)+
  geom_bar(stat="identity")
```
#Gapminder Revisited
```{r}
cp <- read.csv("cell phone total.csv", header = T, row.names=1)
Total <- colSums(cp, na.rm=T)
Average <- colMeans(cp,na.rm = T)
df_y <- as.data.frame(cbind(Total,Average))
df_y$Year <- as.numeric(substr(rownames(df_y),start=2,stop=5))
p1 <- ggplot(aes(x=Year,y=Total),data=df_y)+ geom_point()+
  xlim(1965,2011)
p2 <- ggplot(aes(x=Year,y=Average),data=df_y)+geom_point()+
  xlim(1965,2011)
  
grid.arrange(p1,p2,ncol=1)
```
```{r}
cp <- read.csv("cell phone total.csv", header = T, row.names=1)
df_y <- as.data.frame(colMeans(cp,na.rm = T))
df_y$Year <- as.numeric(substr(rownames(df_y),start=2,stop=5))
ggplot(aes(x=Year,y=Average),data=df_y)+geom_point()+
  xlim(1965,2011)+
  ggtitle('Average/Year of Cell Phone Owned')+
  ggsave('avgpycpplot.jpg')

```


